---
title: "Pre-Processing O3 Data"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(zoo)
library(stats)
```

## Summary

This notebook is almost identical to the previous one, with the exception that the pollutant time series is not O3. The same four cities are taken as examples - London, Berlin, Madrid, and Rome. 

```{r, warning=FALSE, echo=FALSE}
setwd("../data/clean/O3/")

sample.cities <- list()

sample.cities[["London"]] <- read.csv("GB/London/O3_GB_London_2018.csv", row.names = 1)
sample.cities[["Berlin"]] <- read.csv("DE/Berlin/O3_DE_Berlin_2018.csv", row.names = 1)
sample.cities[["Madrid"]] <- read.csv("ES/Madrid/O3_ES_Madrid_2018.csv", row.names = 1)
sample.cities[["Rome"]]   <- read.csv("IT/Roma/O3_IT_Roma_2018.csv", row.names = 1)


sample.cities <- Reduce(
  function(df1, df2) merge(df1, df2, by = "Date", all.x = TRUE, all.y = TRUE), 
  sample.cities
  )

rownames(sample.cities) <- sample.cities[,1]
sample.cities <- sample.cities[,-1]
names(sample.cities) <- c("London", "Berlin", "Madrid", "Rome")

head(sample.cities)
```

## Common trends

A seasonal pattern is clearer for the O3 data, whith concentrations generally peaking in the summer. 

```{r}
par(mfrow = c(2,2))
sample.cities <- sqrt(sample.cities)
for (city in names(sample.cities)) plot.with.dates.axis(sample.cities[[city]], rownames(sample.cities), main = city)
```

The patter remains after taking the square root transformation. 

```{r}
df.plot.with.dates.axis(sample.cities, col = "grey", main = "sample cities after square-root transform")
```

## Removing trends

To remove the seasonal trends I again regress on leading Fourier frequencies and indicators corresponding to weekends. 

```{r}
days <- as.Date(rownames(sample.cities), format = "%Y-%m-%d")
tt <- 1:length(days)

sat <- ifelse(weekdays(days) == "Saturday", 1, 0)
sun <- ifelse(weekdays(days) == "Sunday", 1, 0)

s.qt <- sin(2*pi*tt/365/4)
c.qt <- cos(2*pi*tt/365/4)
s.yr <- sin(2*pi*tt/365)
c.yr <- cos(2*pi*tt/365)

trend.models <- list()

for (city in names(sample.cities)) trend.models[[city]] <- lm(
    sample.cities[[city]] ~ sat + sun + s.qt + c.qt + s.yr + c.yr
    )
```

Visually, the linear model captures the seasonal component in the data well. 

```{r, echo=FALSE}
par(mfrow = c(2,2))
for (city in names(sample.cities))
{
  plot.with.dates.axis(sample.cities[[city]], rownames(sample.cities), main = city)
  lines(trend.models[[city]]$fitted.values, col = "red")
}
```

For most cities the majority of regression coeficients are strongly statistically significant. 

```{r}
summary(trend.models[["London"]])
```

After de-trending visually there does not seem to be a seasonal component. 

```{r, echo=FALSE}
detrended.data <- sample.cities
for (city in names(sample.cities)) detrended.data[[city]] <- trend.models[[city]]$residuals
df.plot.with.dates.axis(detrended.data, col = "grey", main = "sample cities after removing seasonal trends")
```

## Removing serial correlation

Although the seasonal component has been eliminated, the data thus processed is still auto correlated. 

```{r, echo=FALSE}
acf(detrended.data)
```

Again the data may be whitened by fitting an AR model and saving the residuals. 

```{r}
ar.models <- list()
for (city in names(sample.cities)) ar.models[[city]] <- ar(detrended.data[[city]], order.max = 3)
```

Whitening the data in this way seems to remove all series correlation. 

```{r, echo=FALSE}
whitened.data <- sample.cities
for (city in names(sample.cities)) whitened.data[[city]] <- ar.models[[city]]$resid
whitened.data <- whitened.data[-c(1:3),]

acf(whitened.data)
```

Finally, the det-trended and whitened data is plotted below. 

```{r}
df.plot.with.dates.axis(whitened.data, col = "grey", main = "sample cities after trend removal and whitening")
```