---
title: "Multivariate Change point Models for NO2"
output: pdf_document
---

## Summary 

```{r, warning=FALSE}
library(InspectChangepoint)
library(hdbinseg)
library(changepoint.geo)
```

```{r}
source("../R/univaraite_nsp.R")
source("../R/utils.R")
source("../R/pre_process_clean_data.R")
source("../R/multivaraite_nsp/__init__.R")
```


Load all the data... 


```{r}
n.cities <- 400
largest.cities <- read.csv("../data/countries_and_country_codes/largest_cities/largest_cities.csv", 
                           row.names = 1
                           )[,1][1:n.cities]

xx <- stabilize.varaince.detrend(path.to.data = "../data/clean/O3/", 
                                 cities = largest.cities,
                                 target.year = "2020", 
                                 trend.fit.years = "2018", 
                                 show.prog.bar = TRUE
                                 )

```

```{r}
df.plot.with.dates.axis(xx$whitened, main = "some outliers caused by measurement error", col = "grey")
```


```{r}
drop.cities <- c("Portsmouth", "Toulon", "Bratislava")
xx$detrended <- xx$detrended[,!(names(xx$detrended) %in% drop.cities)]
xx$whitened <- xx$whitened[,!(names(xx$whitened) %in% drop.cities)]

df.plot.with.dates.axis(xx$whitened, main = "detrended O3 time series after removing outlier", col = "grey")
```

## Competing methods  

Assuming a pieicewise constsant mean did not seem appropriate for the nuivariate time series. Nevertheless below I apply chanegpoint methods 

```{r}
other.methods <- fit.multivaraute.changepoints(t(xx$whitened))

par(mfrow = c(2,2))

for (ii in 1:length(other.methods))
{
  df.plot.with.dates.axis(xx$whitened, main = names(other.methods)[ii], col = "grey")
  for (jj in other.methods[[ii]]) abline(v = jj, col = "blue")
}
```

Of the three methods that return a sensible number of changepoints there is some disagreement between the changepoint locations. Most European countries implemented a  

```{r}
paste("geometric mapping:", paste(rownames(xx$whitened)[other.methods$`geometric mapping`], collapse = ", "))
paste("double cusum:", paste(rownames(xx$whitened)[other.methods$`double cusum`], collapse = ", "))
paste("sparsified bs:", paste(rownames(xx$whitened)[other.methods$`sparsified binary segmentation`], collapse = ", "))
```

apply to first differenced data...

```{r}
yy <- apply(xx$whitened, 2, diff)
other.methods.diff <- fit.multivaraute.changepoints(t(yy))

par(mfrow = c(2,2))

for (ii in 1:length(other.methods.diff))
{
  df.plot.with.dates.axis(yy, main = names(other.methods.diff)[ii], col = "grey")
  for (jj in other.methods.diff[[ii]]) abline(v = jj, col = "blue")
}
```

## Narrowest Significance Pursuit

```{r}
thresh <- MC.thresh(n = dim(xx$whitened)[2], 
                    tt = dim(xx$whitened)[1], 
                    alpha = 0.05, 
                    K = 500)
```


### NSP with pieiceise constant means

```{r}
m.nsp.pcwsConst <- panel_nsp(t(xx$whitened), thresh = thresh, deg = 0)
```

```{r}
df.plot.with.dates.axis(xx$whitened, col = "grey", main = "NSP with piecewise constant means")
draw_mnsp_rects(m.nsp.pcwsConst$nsp.out, c(-100,100))
```

```{r}
m.nsp.pcwsConst
```
Try isolating only the L2 aggregation

```{r}
m.nsp.pcwsConst.L2 <- panel_nsp(t(xx$whitened), thresh = c(Inf, thresh[2]), deg = 0)
```

```{r}
df.plot.with.dates.axis(xx$whitened, col = "grey", main = "(L2 aggregation only)")
draw_mnsp_rects(m.nsp.pcwsConst.L2$nsp.out, c(-100,100))
```

```{r}
m.nsp.pcwsConst.L2
```


### NSP with piecewise linear means

```{r}
m.nsp.pcwsLin <- panel_nsp(t(xx$whitened), thresh = thresh, deg = 1)
```

```{r}
df.plot.with.dates.axis(xx$whitened, col = "grey", main = "nsp with piecewise linear means")
draw_mnsp_rects(m.nsp.pcwsLin$nsp.out, c(-100,150))
```

```{r}
m.nsp.pcwsLin
```

And again isolating only the L2 aggregations 

```{r}
m.nsp.pcwsLin.L2 <- panel_nsp(t(xx$whitened), thresh = c(Inf, thresh[2]), deg = 1)
```

```{r}
df.plot.with.dates.axis(xx$whitened, col = "grey")
draw_mnsp_rects(m.nsp.pcwsLin.L2$nsp.out, c(-100,150))
```


```{r}
m.nsp.pcwsLin.L2
```

